\TODO{What's the title?}

### 2018-09-18 doc/Paper.tex
Update PDF to a39a0132dc040c4486ad7794acafbe8b27e6ae2c

\usepackage{amsmath}
\usepackage{calc}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{pbox}
\usepackage{tikz}
\usetikzlibrary{arrows,arrows.meta}

@@

\newcommand{\WithMath}[2]{{\parbox{\widthof{#1}}{\centering$#2$}}}

@@

\newcommand{\mimplies}{%
  \hspace{2pt}\makebox[10pt][l]{%
    \tikz[transform canvas={scale=0.5}]{%
    \draw [double distance=3pt,
           line width=0.65pt,
           arrows={Implies[reversed]}-{Implies}] (0,0.17) -- (0.7,0.17);%
    }%
  }%
  \hspace{2pt}%
}

@@

$\mimplies\Rightarrow\implies$

@@

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \mimplies comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
  and has the following signature:
\begin{hscode}\SaveRestoreHook
~~(\mimplies)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\ColumnHook
\end{hscode}\resethooks
  }:
\begin{hscode}\SaveRestoreHook
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~\mimplies~right
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~right
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~left
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~\mimplies~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\ColumnHook
\end{hscode}\resethooks

@@

  We will use repmin\footnote{%
    The repmin problem: Given a tree of integers, replace every integer with the
    minimum integer in the tree, in one pass.%
  } as a running example to illustrate how AGs can be specified using functional
  zippers. The classical solution is the following circular program:
\begin{hscode}\SaveRestoreHook
~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
\\
\texfamily ~~repmin~t~=~t'
\\
\texfamily ~~~{\bfseries where}
\\
\texfamily ~~~~(t',~m')~=~go~t~m'
\\
\texfamily ~~~~go~({\itshape Leaf}~x~~~~~~)~m~=~({\itshape Leaf}~m,~x)
\\
\texfamily ~~~~go~({\itshape Fork}~xs~ys~~)~m~=~({\itshape Fork}~xs'~ys',~min~m\char95 x~m\char95 y)
\\
\texfamily ~~~~~{\bfseries where}
\\
\texfamily ~~~~~~(xs',~m\char95 x)~=~go~xs~m
\\
\texfamily ~~~~~~(ys',~m\char95 y)~=~go~ys~m
\ColumnHook
\end{hscode}\resethooks

\begin{comment}
  Alternatively, we can formulate repmin in terms of an attribute grammar. We
  use \text{\tt localMin} attribute to represent local minimum of a subtree, \text{\tt globalMin}
  to represent the global minimum, and \text{\tt updated} for the solution for a subtree.
\end{comment}

  \TODO{This is a copy-paste} Returning to our example, in
  \cite{viera2009agsfly}, the authors identified three components for solving
  repmin: computing the minimal value, passing down the minimal value from the
  root to the leaves and constructing the resulting tree. In this section, we
  review the Attribute Grammar for repmin that was introduced by
  \cite{swierstra1998designing}, and show how each of the three components
  identified by \cite{viera2009agsfly} in that grammar can be embedded in
  Haskell using our approach.  The attribute grammar for repmin starts by
  defining the underlying data structure, i.e., binary leaf trees. The attribute
  grammar fragments presented in this section follow the standard AG notation of
  \cite{swierstra1998designing}. In this notation, we straightforwardly use the
  \text{\tt Tree~Int} datatype from \TODO{where?}.

  Having defined the structure, we need to define functionality. We start by
  reviewing the AG component that computes the minimal value of a tree:
\begin{hscode}\SaveRestoreHook
~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.localMin~=~@value
\\
\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~@left.localMin~`min`~@right.localMin
\ColumnHook
\end{hscode}\resethooks
  This component declares, using the \text{\tt SYN} keyword, that elements of type \text{\tt Tree}
  synthesize an attribute \text{\tt localMin} of type \text{\tt Int}. Then, a \text{\tt SEM} sentence
  defines how \text{\tt localMin} is computed: when the current tree is a leaf, clearly
  its minimal value is the leaf value itself; when it is the fork of two other
  trees (the \text{\tt left} and the \text{\tt right} subtrees), we compute the minimal values of
  each subtree (i.e., their \text{\tt localMin} attribute), and then their minimal value
  (function \text{\tt min}). In this notation, \text{\tt lhs} refers to the left-hand side symbol
  of the production and @ prefixes a reference to a field.

  Having implemented the first of the three components that solve repmin, we now
  consider the remaining two. We start by implementing the construction of the
  result of repmin, a tree with all leaves being the minimum of the original
  one.
\begin{hscode}\SaveRestoreHook
~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
\\
\texfamily ~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated~@right.updated
\ColumnHook
\end{hscode}\resethooks
  The implementation of sres places in each leaf of a tree the value of the ival
  attribute. This value corresponds to the minimal value of the global tree,
  that still needs to be passed down to all the nodes in the tree. This
  corresponds exactly to the third component that we still need to implement. In
  order to bind the minimal value being computed (attribute smin) with the
  minimal value that is passed down through the tree (attribute ival), it is
  common, in the AG setting, to add a new data-type definition,

@@

  Repmin as an AG...

>>>>

\usepackage{amsmath}
\usepackage{calc}
\usepackage{csquotes}
\usepackage{color}
\usepackage{comment}
\usepackage{float}
\usepackage{graphicx}
\usepackage{url}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage{pbox}

@@

\newcommand{\WithMath}[2]{{\parbox[][][b]{\widthof{#1}}{\centering$#2$}}}

% For the circular variant of repmin

% For the specification of AGs

@@

@@

@@

  \newpage
  using which we can perform multiple updates efficiently and with minimal code
  bloat\footnote{%
  Operator \text{\tt \char62{}\char61{}\char62{}} comes from \text{\tt Control\char46{}Monad} module in \text{\tt base}
  and has the following signature:
\begin{hscode}\SaveRestoreHook
~~(>=>)~::~{\itshape Monad}~m~=>~(a~\WithMath{->}{\rightarrow}~m~b)~\WithMath{->}{\rightarrow}~(b~\WithMath{->}{\rightarrow}~m~c)~\WithMath{->}{\rightarrow}~a~\WithMath{->}{\rightarrow}~m~c
\ColumnHook
\end{hscode}\resethooks
  }:
\begin{hscode}\SaveRestoreHook
~~modifyExample~::~{\itshape IO}~()
\\
\texfamily ~~modifyExample~=~print~\$
\\
\texfamily ~~~~lzEnter~@{\itshape Int}~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~right
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(const~[])
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~(+1))
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~left
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~lzModify~(return~\WithMath{.}{\circ}~negate)
\\
\texfamily ~~~~~~~~~~~~~~~~~>=>~return~\WithMath{.}{\circ}~lzLeave~\$
\\
\texfamily ~~~~~~[1,~2,~3,~4,~5]
\ColumnHook
\end{hscode}\resethooks

@@

  Let us consider the repmin\footnote{%
    The repmin problem:
    \begin{displayquote}
      Given a tree of integers, replace every integer with the
      minimum integer in the tree, in one pass.
    \end{displayquote}
  } problem as an example of a problem that requires multiple traversals. The
  classical solution is the following circular program:
\begin{hscode}\SaveRestoreHook
~~repmin~::~{\itshape Tree}~{\itshape Int}~\WithMath{->}{\rightarrow}~{\itshape Tree}~{\itshape Int}
\\
\texfamily ~~repmin~t~=~t'
\\
\texfamily ~~~~{\bfseries where}~(t',~m')~=~go~t~m'
\\
\texfamily ~~~~~~~~~~go~({\itshape Leaf}~x~~~~)~m~=~({\itshape Leaf}~m,~x)
\\
\texfamily ~~~~~~~~~~go~({\itshape Fork}~xs~ys)~m~=~({\itshape Fork}~xs'~ys',~min~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$}~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})
\\
\texfamily ~~~~~~~~~~~~{\bfseries where}~(xs',~\parbox[][3.5pt][t]{\widthof{mx}}{m$_{\text{x}}$})~=~go~xs~m
\\
\texfamily ~~~~~~~~~~~~~~~~~~(ys',~\parbox[][3.5pt][t]{\widthof{my}}{m$_{\text{y}}$})~=~go~ys~m
\ColumnHook
\end{hscode}\resethooks

  Although quite elegant, the code lacks modularity and is very difficult to
  reason about. Attribute Grammars provide a more modular approach. Viera et
  al\cite{viera2009agsfly} identified three steps for solving repmin: computing
  the minimal value, passing it down from the root to the leaves, and
  constructing the resulting tree. We can associate each step with an
  attribute\cite{swierstra1998designing}:
  \begin{itemize}
  \item A synthesized attribute \text{\texfamily localMin\Sp ::\Sp {\itshape Int}} represents the minimum value
        of a subtree. Computing the minimal value thus corresponds to evaluation
        of the \text{\texfamily localMin} attribute for the root tree.
  \item An inherited attribute \text{\texfamily globalMin\Sp ::\Sp {\itshape Int}} is used to pass down the
        minimal value.
  \item Finally, a synthesized attribute \text{\texfamily updated\Sp ::\Sp {\itshape Tree}\Sp {\itshape Int}} is the subtree
        with leaf values replaced by values of their \text{\texfamily globalMin} attributes. The
        solution is thus the value of \text{\texfamily updated} attribute for the root tree.
  \end{itemize}
  The obtained AG is presented in figure~\ref{fig:repmin-AG}. \TODO{Do we
  actually need to explain the algorithms here? It seems a little bit childish
  to explain how to compute the minimum of a binary tree... If we absolutely
  have to explain stuff, maybe just put it in the caption.}

\begin{figure}
\centering
\fbox{\parbox{\linewidth}{%
\begin{hscode}\SaveRestoreHook
~~SYN~{\itshape Tree}~{\itshape Int}~[localMin~:~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.localMin~=~@value
\\
\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.localMin~=~min~@left.localMin
\\
\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.localMin
\\
\texfamily 
\\
\texfamily ~~SYN~{\itshape Tree}~{\itshape Int}~[updated~:~{\itshape Tree}~{\itshape Int}]
\\
\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Leaf}~lhs.updated~=~{\itshape Leaf}~@lhs.globalMin
\\
\texfamily ~~~~~~~~~~~~~~~|~{\itshape Fork}~lhs.updated~=~{\itshape Fork}~@left.updated
\\
\texfamily ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~@right.updated
\\
\texfamily 
\\
\texfamily 
\\
\texfamily ~~INH~{\itshape Tree}~{\itshape Int}~[~globalMin~:~{\itshape Int}~]
\\
\texfamily ~~SEM~{\itshape Tree}~{\itshape Int}~|~{\itshape Fork}~left.globalMin~~=~@lhs.globalMin
\\
\texfamily ~~~~~~~~~~~~~~~~~~~~~~right.globalMin~=~@lhs.globalMin
\\
\texfamily 
\\
\texfamily ~~DATA~{\itshape Root}~|~{\itshape Root}~tree~:~{\itshape Tree}~{\itshape Int}
\\
\texfamily ~~SEM~~{\itshape Root}~|~{\itshape Root}~tree.globalMin~=~@tree.localMin
\ColumnHook
\end{hscode}\resethooks
%
}}
\caption{%
  Attribute grammar for repmin. The syntax is closely mirrors the one used
  in\cite{swierstra1998designing}. \text{\texfamily SYN} and \text{\texfamily INH} introduce synthesized
  and inherited attributes respectively. \text{\texfamily SEM} is used for defining semantic
  rules. A new data type \text{\texfamily {\itshape Root}} is introduced as it is common in the AG setting
  to ``connect'' \text{\texfamily localMin} with \text{\texfamily globalMin}.
}
\label{fig:repmin-AG}
\end{figure}

  We now move on to embed this attribute grammar into Haskell. Semantic rules
  simply become functions, which, given a zipper, return values of the
  attributes. For example,
\begin{hscode}\SaveRestoreHook
~~localMin~::~{\itshape {\itshape Lib}.Zipper}~{\itshape Cxt}~{\itshape Attributes}~({\itshape Tree}~{\itshape Int})~\WithMath{->}{\rightarrow}~{\itshape Int}
\\
\texfamily ~~localMin~z@({\itshape {\itshape Lib}.Zipper}~hole)~=~{\bfseries case}~whereami~hole~{\bfseries of}
\\
\texfamily ~~~~{\itshape C\char95 Leaf}~\WithMath{->}{\rightarrow}~{\bfseries let}~({\itshape Leaf}~x)~=~hole~{\bfseries in}~x
\\
\texfamily ~~~~{\itshape C\char95 Fork}~\WithMath{->}{\rightarrow}~min~(localMin~\$~{\itshape Lib}.unsafeChild~0~z)
\\
\texfamily ~~~~~~~~~~~~~~~~~~(localMin~\$~{\itshape Lib}.unsafeChild~1~z)
\ColumnHook
\end{hscode}\resethooks

@@
